---
title: "gene-plot-Pi-shiny-beta"
author: "Hubert Kicinski"
date: "2024-10-09"
output: 
  html_notebook:
    toc: false
    toc_float: false
    code_folding: hide
runtime: shiny
---
<style type="text/css">
.main-container {
  max-width: 100%;
  margin-left: auto;
  margin-right: auto;
}
</style>
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```
```{r load_data_library, echo=FALSE}
load("R_obj_updated.RData")
# install the package if not already available
suppressMessages(library(tidyverse))
suppressMessages(library(data.table))
suppressMessages(library(cowplot))
suppressMessages(library(shiny))
```
```{r plot_gene, echo=FALSE}
myGenePlot <- function(lc = lcpm, gene = "CAGL0B02475g", sample_info = sample_info) {
  if(!(gene %in% rownames(lc))) {
    return(ggplot() + 
             annotate("text", x = 0.5, y = 0.5, label = "Gene ID not found in expression data") + 
             theme_void())
  }
  
  exn <- lc[gene,]
  
  dt <- data.table(
    Sample = colnames(lc),
    Timepoint = sample_info$Timepoint,
    Replicate = sample_info$Replicate,
    exn = as.numeric(exn)
  )
  
  time_order <- c("0min", "15min", "30min", "45min", "1h", "1.5h", "2h", "3h", "4h", "6h", "8h")
  dt$Timepoint <- factor(dt$Timepoint, levels = time_order)
  
  p <- ggplot(dt, aes(x = Timepoint, y = exn, color = factor(Replicate), group = Replicate)) +
    geom_point() + 
    geom_line() +
    ylab("log2 count per million") + 
    ggtitle(paste("Expression of", gene)) +
    theme_cowplot() + 
    theme(axis.text.x = element_text(angle = 45, size = rel(0.75), hjust = 1)) +
    scale_color_discrete(name = "Replicate")
  
  p + panel_border() + background_grid()
}
```
```{r shiny_app, echo=FALSE}
# Updated search_gene function
search_gene <- function(query) {
  query <- toupper(query)
  
  # Check for CAGL0 ID directly in LCPM rownames
  if (grepl("^CAGL0", query) && query %in% rownames(lcpm)) {
    return(query)
  }
  
  # Check for GWK60 ID and map to CAGL0 ID
  if (grepl("^GWK60_", query)) {
    cagl_id <- anno$GeneID[anno$GWK60_ID == query]
    if (length(cagl_id) == 1 && cagl_id %in% rownames(lcpm)) {
      return(cagl_id)
    }
  }
  
  # Check for gene name and map to CAGL0 ID
  if (query %in% anno$GeneName) {
    cagl_id <- anno$GeneID[anno$GeneName == query]
    if (length(cagl_id) == 1 && cagl_id[1] %in% rownames(lcpm)) {
      return(cagl_id)
    }
  }
  
  # Check for S. cerevisiae name and map to CAGL0 ID
  if (query %in% cgToSc$scname) {
    cagl_id <- cgToSc$cgid[cgToSc$scname == query]
    if (length(cagl_id) == 1 && cagl_id[1] %in% rownames(lcpm)) {
      return(cagl_id)
    }
  }
  
  # If no exact match is found, try to partially match
  partial_matches <- c(
    rownames(lcpm)[grepl(query, rownames(lcpm), ignore.case = TRUE)],
    anno$GeneID[grepl(query, anno$GeneID, ignore.case = TRUE) | 
                grepl(query, anno$GeneName, ignore.case = TRUE) | 
                grepl(query, anno$GWK60_ID, ignore.case = TRUE)],
    cgToSc$cgid[grepl(query, cgToSc$scname, ignore.case = TRUE)]
  )
  
  partial_matches <- unique(partial_matches)
  
  if (length(partial_matches) == 1) {
    return(partial_matches[1])  # Return the first partial match
  }
  
  return(NULL)
}
library(shiny)
library(ggplot2)
library(cowplot)
ui <- fillPage(
  
  navbarPage(
  " ",   
  navbarMenu("species", 
    tabPanel("Candida glabrata", 
  tags$head(
    tags$style(HTML("
      body, html {
        height: 100%;
        width: 100%;
        margin: 0;
        padding: 0;
      }
      #sidebar {
        width: 300px;
        padding: 20px;
        background-color: #f5f5f5;
        height: 100vh;
        position: fixed;
        left: 0;
        top: 0;
        overflow-y: auto;
        border-right: 1px solid #ddd;
      }
      #mainPanel {
        margin-left: 300px;
        height: 100vh;
        display: grid;
        grid-template-rows: auto 1fr;
        gap: 20px;
        padding: 20px;
        overflow-y: auto;
      }
      .plot-container {
        background: white;
        border: 1px solid #ddd;
        border-radius: 4px;
        padding: 15px;
      }
      .results-container {
        display: grid;
        grid-template-columns: 1fr 1fr;
        gap: 20px;
        align-content: start;
      }
      .results-table {
        background: white;
        border: 1px solid #ddd;
        border-radius: 4px;
        padding: 15px;
      }
      .table-caption {
        font-weight: bold;
        margin-bottom: 10px;
        color: #333;
      }
      #gene_info {
        background: white;
        border: 1px solid #ddd;
        border-radius: 4px;
        padding: 15px;
        margin-top: 20px;
        white-space: pre-wrap;
      }
      #plot_button {
        width: 100%;
        margin-top: 10px;
        padding: 8px;
        background-color: #337ab7;
        color: white;
        border: none;
        border-radius: 4px;
        cursor: pointer;
      }
      #plot_button:hover {
        background-color: #286090;
      }
      .shiny-input-container {
        width: 100%;
      }
    "))
  ),
  
  div(id = "sidebar",
    h3("Gene Expression Analysis", style = "margin-top: 0;"),
    textInput("genename", "Enter gene name or ID:", "Pho4"),
    actionButton("plot_button", "Generate Plot"),
    hr(),
    verbatimTextOutput("gene_info")
  ),
  
  div(id = "mainPanel",
    div(class = "plot-container",
      plotOutput("gene_plot")
    ),
    div(class = "results-container",
      div(class = "results-table",
        div(class = "table-caption", "C. glabrata Search Results"),
        tableOutput("search_results_cg")
      ),
      div(class = "results-table",
        div(class = "table-caption", "S. cerevisiae Search Results"),
        tableOutput("search_results_sc")
      )
    )
  )
),
    tabPanel("Saccharomyces cerevisiae", "place holder 1"),
    tabPanel("Kluyveromyces lactis", "place holder 2"),
    tabPanel("Candida albicans", "place holder 3")
  )
))
server <- function(input, output, session) {
  selected_gene <- reactiveVal()
  
  observeEvent(input$plot_button, {
    gene_id <- search_gene(input$genename)
    selected_gene(gene_id)
  })
  
  # Render the plot based on the selected gene
  output$gene_plot <- renderPlot({
    req(selected_gene())
    gene <- selected_gene()
    if (!is.null(gene) && gene %in% rownames(lcpm)) {
      myGenePlot(gene = gene, lc = lcpm, sample_info = sample_info)
    } else {
      ggplot() + 
        annotate("text", x = 0.5, y = 0.5, label = "Gene not found or invalid input") + 
        theme_void()
    }
  })
  
  # Render the gene information in the sidebar
  output$gene_info <- renderText({
    req(selected_gene())
    gene <- selected_gene()
    if (!is.null(gene)) {
      gene_info <- anno[anno$GeneID == gene, ]
      if (nrow(gene_info) > 0) {
        paste("Gene ID:", gene,
              "\nGene Name:", gene_info$GeneName,
              "\nGWK60 ID:", gene_info$GWK60_ID,
              "\nFound in dataset:", if(gene %in% rownames(lcpm)) "Yes" else "No")
      } else {
        paste("Gene ID:", gene, "\nFound in dataset:", if(gene %in% rownames(lcpm)) "Yes" else "No")
      }
    } else {
      "No gene found matching the input."
    }
  })
  
  # Display search results for C. glabrata
  output$search_results_cg <- renderTable({
    req(input$genename)
    query <- toupper(input$genename)
    anno %>% 
      filter(GeneID == query | GeneName == query | GWK60_ID == query |
               grepl(query, GeneID, ignore.case = TRUE) | 
               grepl(query, GeneName, ignore.case = TRUE) |
               grepl(query, GWK60_ID, ignore.case = TRUE)) %>% 
      select(GeneID, GeneName, GWK60_ID) %>%
      head()
  }, caption = "C. glabrata search results")
  
  # Display search results for S. cerevisiae
  output$search_results_sc <- renderTable({
    req(input$genename)
    query <- toupper(input$genename)
    cgToSc %>% 
      filter(cgid == query | scname == query |
               grepl(query, cgid, ignore.case = TRUE) | 
               grepl(query, scname, ignore.case = TRUE)) %>% 
      head()
  }, caption = "S. cerevisiae search results")
}
shinyApp(ui = ui, server = server, options = list(height = 1080))